---
title: "Homework 5"
author: "Juntao Zhang"
date: "06/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read and inspect data set
```{r read data}
data <-read.csv("C:/Users/joann/OneDrive/Desktop/missing data/week 2/aug_train.csv", 
                 na.strings = "")

```

#Encode character variables

```{r encode}
unique(data$relevent_experience )
library(plyr)
data$relevent_experience <- revalue(data$relevent_experience, 
                                    c("Has relevent experience"=1))
data$relevent_experience <- revalue(data$relevent_experience, 
                                    c("No relevent experience"=0))
data$relevent_experience <-as.numeric(data$relevent_experience)

unique(data$last_new_job)
data$last_new_job <- revalue(data$last_new_job, c("never"=0))
data$last_new_job <- revalue(data$last_new_job, c(">4"=5))
data$last_new_job <-as.numeric(data$last_new_job)

unique(data$enrolled_university )
data$enrolled_university <- revalue(data$enrolled_university, 
                                    c("no_enrollment"=0))
data$enrolled_university <- revalue(data$enrolled_university, 
                                    c("Part time course"=1))
data$enrolled_university <- revalue(data$enrolled_university,
                                    c("Full time course" = 2))
data$enrolled_university <-as.numeric(data$enrolled_university)

unique(data$education_level)
data$education_level <- as.numeric(factor(data$education_level, 
                                          levels = c("Primary School",
                                                     "High School","Graduate",
                                                     "Masters","Phd")))
unique(data$gender) 
data$gender <- as.factor(data$gender)

```

```{r select data}
#I will keep the variables that can be used for my analysis
library(dplyr)
data2 = select(data,'training_hours','gender','relevent_experience',
               'last_new_job','enrolled_university','education_level','target')

```


```{r generate missing values}
#Generate missing values for training_hours depending on one variable
library(dplyr)
data_new = select(data,'city_development_index','training_hours')
library(mice)
cont_cat = ampute(data_new,prop = 0.2,patterns=c(1,0),mech = "MAR")$amp
data2['training_hours'] = cont_cat['training_hours']


#check again for the generated missing values
sapply(data2, function(x) sum(is.na(x)))
```


```{r imputation}
#Q1

library(mi)
#Run mi with 5 chains and 50 iterations on the dataset

# Create the missing data frame object
mdf = missing_data.frame(data2)

# Examine the default settings
show(mdf)

# Running the chains
imputations <- mi(mdf, n.chains = 5, n.iter=50,max.minutes = 20)

```


```{r check convergence}
#Q2

#Check convergence/diagnostics and make changes if necessary
Rhats(imputations)
round(mipply(imputations, mean, to.matrix = TRUE), 3)

```

#make changes for last_new_job and training hours
#since they have unequal means for each chain 
#the inspected problems are on the type of these two variables
#training hours are always >0 and last new job is a ordered categorical variable
```{r make change}

mdf <- change(mdf, y = "last_new_job", what = "type", 
              to = "ordered-categorical")
mdf <- change(mdf, y = "training_hours", what = "type", to = "pos")
show(mdf)
```

```{r check again}

# Rerunning the chains
imputations <- mi(mdf, n.chains = 5, n.iter=50)
round(mipply(imputations, mean, to.matrix = TRUE), 3)
converged <- mi2BUGS(imputations)
Rhats(imputations)
mean_th = converged[, , 1]
# Traceplot of mean imputed training hours
ts.plot(mean_th[,1], col=1)
lines(mean_th[,2], col= 2)
lines(mean_th[,3], col= 3)
lines(mean_th [,4], col= 4)
lines(mean_th [,5], col= 5)

mean_l = converged[, , 4]
# Traceplot of mean imputed last new job
ts.plot(mean_l[,1], col=1)
lines(mean_l[,2], col= 2)
lines(mean_l[,3], col= 3)
lines(mean_l [,4], col= 4)
lines(mean_l [,5], col= 5)

#check for the plots 
plot(imputations)

#the results converged 

```

```{r pool}

#Q3

#Pool the results and report the estimated equation
analysis <- pool(target ~ training_hours+gender+relevent_experience+
                   last_new_job+enrolled_university+education_level,imputations)
display(analysis)

#the estimated equation of the pooled result is as shown above 

```

#Q4
#compare pooled result with complete dataset result

```{r compare}
#original complete data set
data_complete = na.omit(data)

#Linear regression analysis for the target variable
#0-not looking for a job change 1-looking for a job change


#model with the original complete data cases
model2 = lm(target ~ training_hours+gender+relevent_experience+last_new_job+
              enrolled_university+education_level, data = data_complete)
summary(model2)

```


#Comparing the two results:
#The estimated coefficients for the original complete data set are all very small but for the pooled result from the imputed data set, relevant experience, enrolled university and education level are showing a larger effect (estimated coefficient) for target. 
